---
title: "BibTeX and CFF"
subtitle: A potential crosswalk
bibliography: REFERENCES.bib
author: Diego Hernangómez
description: >-
  This article presents a crosswalk between BibTeX and Citation File Format 
  [@druskat_citation_2021], as it is performed by the cffr package 
  [@hernangomez2021].
abstract: >-
  This article introduces a crosswalk between **BibTeX** and the Citation File 
  Format (CFF) [@druskat_citation_2021], as implemented by the **cffr** package 
  [@hernangomez2021]. The crosswalk aims to facilitate seamless translation 
  between these two reference formats. Specifically, it proposes various 
  crosswalk models tailored to different **BibTeX** entry types 
  [@patashnik1988]. Additionally, the article includes practical examples using 
  real **BibTeX** entries and offers tips for developers interested in 
  implementing this crosswalk across various programming languages.
link-citations: yes
documentclass: article
editor_options:
  markdown:
    wrap: 150
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{BibTeX and CFF}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ""
)

options(width = 60)
```

## Citation

Please cite this article as:

```{r cit, echo=FALSE, results='asis'}
thisart <- bibentry("article",
  title = "{BibTeX} and {CFF}, a potential crosswalk",
  key = "hernangomez2022",
  author = "Diego Hernangómez",
  journal = "The {cffr} package",
  year = 2022,
  volume = "Vignettes",
  doi = "10.21105/joss.03900",
  url = "https://docs.ropensci.org/cffr/articles/bibtex_cff.html"
)
cat("  \n")
thisart
```

A **BibTeX** entry for LaTeX users:

``` bibtex
@article{hernangomez2022,
    title        = {{BibTeX} and {CFF}, a potential crosswalk},
    author       = {Diego Hernangómez},
    year         = 2022,
    journal      = {The {cffr} package},
    volume       = {Vignettes},
    doi          = {10.21105/joss.03900},
    url          = {https://docs.ropensci.org/cffr/articles/bibtex_cff.html}
}
```

## BibTeX and R

[**BibTeX**](https://en.wikipedia.org/wiki/BibTeX) is a well-known format for storing references created by [Oren
Patashnik](https://en.wikipedia.org/wiki/Oren_Patashnik "Oren Patashnik") and [Leslie
Lamport](https://en.wikipedia.org/wiki/Leslie_Lamport "Leslie Lamport") back in 1985. **BibTeX** can be reused by other software, such as
[LaTeX](https://en.wikipedia.org/wiki/LaTeX), to add references to scholarly works. An example structure of a **BibTeX** entry would be:

``` bibtex
@book{einstein1921,
    title        = {Relativity: The Special and the General Theory},
    author       = {Einstein, A.},
    year         = 1920,
    publisher    = {Henry Holt and Company},
    address      = {London, United Kingdom},
    isbn         = 9781587340925
}
```

In this case, the entry (identified as `einstein1921`) refers to a book. This entry can then be used in a document to include references to it. In
**R** [@R_2021], we can replicate this structure using the `bibentry()` and `toBibtex()` functions:

```{r bibentry, comment="#>"}
entry <- bibentry("book",
  key = "einstein1921",
  title = "Relativity: The Special and the General Theory",
  author = person("A.", "Einstein"),
  year = 1920,
  publisher = "Henry Holt and Company",
  address = "London, United Kingdom",
  isbn = 9781587340925,
)

toBibtex(entry)
```

The final results of the entry as a text string would be parsed as[^1]:

[^1]: By default R Pandoc would generate the cite on the Chicago author-date format [@rmarkdowncookbook2020]

```{r echo=FALSE, results='asis'}
entry
```

Additionally, the **cffr** package [@hernangomez2021] incorporates a function `cff_from_bibtex()` that can be used to read and transform **BibTeX**
strings into different formats:

```{r cffbibread,comment="#>"}
string <- "@book{einstein1921,
    title        = {Relativity: The Special and the General Theory},
    author       = {Einstein, A.},
    year         = 1920,
    publisher    = {Henry Holt and Company},
    address      = {London, United Kingdom},
    isbn         = 9781587340925}"

# To cff
library(cffr)
cff_format <- cff_from_bibtex(string)

cff_format

# To citation R format and bibtex
citation_format <- cff_to_bibtex(cff_format)
class(citation_format)
citation_format

toBibtex(citation_format)
```

## BibTeX Definitions

@patashnik1988 provides a comprehensive explanation of the **BibTeX** formats. Let's distinguish between **Entries** and **Fields**.

### Entries {#entries}

Each entry type defines a different type of work. The 14 entry types defined in **BibTeX**[^2] are:

[^2]: Other implementations similar to **BibTeX**, as [BibLaTeX](https://www.ctan.org/pkg/biblatex), expand the definitions of entries including other
    types as **online**, **software** or **dataset**. On **BibTeX** these entries should be reclassified to **misc**.

1.  **\@article**: An article from a journal or magazine.
2.  **\@book**: A book with an explicit publisher.
3.  **\@booklet**: A work that is printed and bound, but without a named publisher or sponsoring institution.
4.  **\@conference**: Equivalent to **\@inproceedings**, included for [Scribe](https://en.wikipedia.org/wiki/Scribe_(markup_language)) compatibility.
5.  **\@inbook**: A part of a book, which may be a chapter (or section) and/or a range of pages.
6.  **\@incollection**: A part of a book having its own title.
7.  **\@inproceedings**: An article in conference proceedings.
8.  **\@manual**: Technical documentation.
9.  **\@mastersthesis**: A Master's thesis.
10. **\@misc**: Use this type when nothing else fits.
11. **\@phdthesis**: A PhD thesis.
12. **\@proceedings**: The proceedings of a conference.
13. **\@techreport**: A report published by a school or other institution, usually numbered within a series.
14. **\@unpublished**: A document having an author and title, but not formally published.

Regarding the entries, the `bibentry()` R function does not implement **\@conference**. However, we can replace that key with **\@inproceedings**,
given that the definition is identical.

### Fields {#fields}

Similar to the **Entries**, @patashnik1988 also provides definitions for each of the possible standard **BibTeX fields**[^3]. While an entry must
include certain required fields, it can also include additional fields that might be ignored in the raw implementation of **BibTeX**. Let's explore
some of these fields:

[^3]: As in the case of the entries, other implementations based on **BibTeX** may recognize additional fields.

1.  **address**: Typically represents the address of the **publisher** or another **institution**.
2.  **annote**: An annotation. Although not used by standard bibliography styles, it may be relevant for producing annotated bibliographies.
3.  **author**: Contains the name(s) of the author(s), following the format described in the LaTeX book [@lamport86latex].
4.  **booktitle**: Refers to the title of a book, part of which is being cited. For **\@book** entries, use the **title** field instead.
5.  **chapter**: Indicates a chapter (or section) number.
6.  **crossref**: Refers to the database key of the entry being cross-referenced.
7.  **edition**: Specifies the edition of a **\@book**, e.g., `"Second"`. The ordinal should have the first letter capitalized; standard styles
    convert to lowercase when necessary.
8.  **editor**: Contains the name(s) of editor(s), following the conventions in the LaTeX book [@lamport86latex]. If there is also an **author**
    field, the **editor** field specifies the editor of the book or collection where the reference appears.
9.  **howpublished**: Describes how something unusual has been published. The first word should be capitalized.
10. **institution**: Represents the sponsoring institution of a technical report.
11. **journal**: Refers to the name of a journal.
12. **key**: Used for alphabetizing, cross-referencing, and creating a label when **author** information is missing.
13. **month**: Indicates the month in which the work was published or, for an unpublished work, when it was written. Use the standard three-letter
    abbreviation (e.g., `jan`, `feb`, `mar`) as described in Appendix B.1.3 of the LaTeX book [@lamport86latex].
14. **note**: Provides any additional information that can assist the reader. The first word should be capitalized.
15. **number**: Represents the number of a journal, magazine, technical report, or a work in a series. An issue of a journal or magazine is usually
    identified by its **volume** and number. Organizations issuing technical reports often assign them numbers, and sometimes books are given numbers
    within a named series.
16. **organization**: Refers to the organization that sponsors a **\@conference** or publishes a manual.
17. **pages**: Specifies one or more page numbers or a range of numbers (e.g., `42--111` or `7,41,73--97` or `43+`).
18. **publisher**: Indicates the publisher's name.
19. **school**: Provides the name of the school where a thesis was written.
20. **series**: Specifies the name of a series or set of books. When citing an entire book, the **title** field gives its title, and an optional
    **series** field provides the name of a series or multi-volume set in which the book is published.
21. **title**: Represents the work's title.
22. **type**: Describes the type of a technical report (e.g., "Research Note").
23. **volume**: Refers to the volume of a journal or multi-volume book.
24. **year**: Indicates the year of publication or, for an unpublished work, the year it was written. Generally, it should consist of four numerals
    (e.g., `1984`).

In **BibTeX**, there exists a strict relationship between **Entries** and **Fields**. Depending on the type of entry, certain fields are required,
while others are optional or even ignored.

The following table summarizes the relationship between **Entries** and **Fields** in **BibTeX**. Required fields are flagged as **x**, and optional
fields are flagged as **o**. Fields enclosed in parentheses (**(x), (o)**) indicate varying degrees of flexibility regarding the field requirements.
For more detailed information, refer to @patashnik1988.

```{r entry_fields1, echo=FALSE}
df_ini <- data.frame(
  stringsAsFactors = FALSE,
  field = c(
    "title", "year",
    "author", "note", "month", "address",
    "publisher", "editor", "number", "volume", "pages",
    "series", "booktitle", "edition", "type", "chapter",
    "organization", "howpublished", "institution",
    "journal", "school", "annote", "crossref", "key"
  ),
  article = c(
    "**x**",
    "**x**", "**x**", "**o**", "**o**", NA, NA, NA, "**o**",
    "**o**", "**o**", NA, NA, NA, NA, NA, NA, NA,
    NA, "**x**", NA, NA, NA, NA
  ),
  book = c(
    "**x**",
    "**x**", "**(x)**", "**o**", "**o**", "**o**", "**x**",
    "**(x)**", "**(o)**", "**(o)**", NA, "**o**", NA,
    "**o**", NA, NA, NA, NA, NA, NA, NA, NA, NA, NA
  ),
  booklet = c(
    "**x**",
    "**o**", "**o**", "**o**", "**o**", "**o**", NA, NA, NA,
    NA, NA, NA, NA, NA, NA, NA, NA, "**o**", NA, NA,
    NA, NA, NA, NA
  ),
  inbook = c(
    "**x**",
    "**x**", "**(x)**", "**o**", "**o**", "**o**", "**x**",
    "**(x)**", "**(o)**", "**(o)**", "**(x)**",
    "**o**", NA, "**o**", "**o**", "**(x)**", NA, NA, NA,
    NA, NA, NA, NA, NA
  ),
  incollection = c(
    "**x**",
    "**x**", "**x**", "**o**", "**o**", "**o**", "**x**",
    "**o**", "**(o)**", "**(o)**", "**o**", "**o**",
    "**x**", "**o**", "**o**", "**o**", NA, NA, NA, NA,
    NA, NA, NA, NA
  ),
  conference.inproceedings = c(
    "**x**",
    "**x**", "**x**", "**o**", "**o**", "**o**", "**o**",
    "**o**", "**(o)**", "**(o)**", "**o**", "**o**",
    "**x**", NA, NA, NA, "**o**", NA, NA, NA, NA, NA,
    NA, NA
  ),
  manual = c(
    "**x**",
    "**o**", "**o**", "**o**", "**o**", "**o**", NA, NA, NA,
    NA, NA, NA, NA, "**o**", NA, NA, "**o**", NA,
    NA, NA, NA, NA, NA, NA
  ),
  mastersthesis.phdthesis = c(
    "**x**",
    "**x**", "**x**", "**o**", "**o**", "**o**", NA, NA, NA,
    NA, NA, NA, NA, NA, "**o**", NA, NA, NA, NA, NA,
    "**x**", NA, NA, NA
  ),
  misc = c(
    "**o**",
    "**o**", "**o**", "**o**", "**o**", NA, NA, NA, NA, NA,
    NA, NA, NA, NA, NA, NA, NA, "**o**", NA, NA, NA,
    NA, NA, NA
  ),
  proceedings = c(
    "**x**",
    "**x**", NA, "**o**", "**o**", "**o**", "**o**",
    "**o**", "**(o)**", "**(o)**", NA, "**o**", NA, NA, NA,
    NA, "**o**", NA, NA, NA, NA, NA, NA, NA
  ),
  techreport = c(
    "**x**",
    "**x**", "**x**", "**o**", "**o**", "**o**", NA, NA,
    "**o**", NA, NA, NA, NA, NA, "**o**", NA, NA, NA,
    "**x**", NA, NA, NA, NA, NA
  ),
  unpublished = c(
    "**x**",
    "**o**", "**x**", "**x**", "**o**", NA, NA, NA, NA, NA,
    NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA,
    NA, NA
  )
)

df_ini[is.na(df_ini)] <- ""
df_ini$field <- paste0("**", df_ini$field, "**")

getnames <- names(df_ini)
getnames <- gsub(".", ", \\@", getnames, fixed = TRUE)
getnames <- paste0("**\\@", getnames, "**")


t1 <- df_ini[, c(1:7)]
nm1 <- getnames[1:7]

knitr::kable(t1,
  col.names = nm1,
  align = c("l", rep("c", 6)),
  caption = "**BibTeX**, required fields by entry"
)
```

```{r entry_fields2, echo=FALSE}
t2 <- df_ini[, c(1, 8:13)]
nm2 <- getnames[c(1, 8:13)]
knitr::kable(t2,
  col.names = nm2,
  align = c("l", rep("c", 6)),
  caption = "(cont) **BibTeX**, required fields by entry"
)
```

It can be seen that only a subset of fields is necessary for any entry. For instance, **title**, **year**, and **author** are either required or
optional in almost every entry, whereas **crossref**, **annote**, or **key** are never mandatory.

## Citation File Format

[Citation File Format (CFF](https://citation-file-format.github.io/)) [@druskat_citation_2021] consists of plain text files containing both human- and
machine-readable citation information for software and datasets. Within [CFF]{.underline}, there are two important keys: `preferred-citation` and
`references`, which play a crucial role in citing and referring to related works:

-   **`preferred-citation`**: Refers to another work that should be cited instead of the software or dataset itself.
-   **`references`**: Includes reference(s) to other creative works related to the software or dataset. Similar to a list of references in a scholarly
    paper, these references may encompass papers describing the abstract concepts of the software or algorithms implemented in the software version.

These two keys are expected to be `definition.reference objects`, as defined in the [Guide to Citation File Format schema version
1.2.0](https://github.com/citation-file-format/citation-file-format/blob/main/schema-guide.md#preferred-citation), and they may contain the following
keys:

```{r refkeys, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
library(cffr)

# Fill with whites
init <- paste0("[", cff_schema_definitions_refs(), "]{.underline}")

l <- c(init, rep("", 4))


refkeys <- matrix(l, ncol = 5, byrow = TRUE)

knitr::kable(refkeys,
  caption = "Valid keys on [CFF]{underline} `definition-reference` objects"
)
```

These keys are equivalent to the **BibTeX** [fields](#fields), with the exception of the key **type**. In [CFF]{.underline}, this key defines the type
of work[^4], making it analogous to the **BibTeX** [entries](#entries).

[^4]: See a complete list of possible values of [CFF]{.underline} type on the [Guide to Citation File Format schema version
    1.2.0](https://github.com/citation-file-format/citation-file-format/blob/main/schema-guide.md#definitionsreferencetype).

## Proposed Crosswalk

The **cffr** package [@hernangomez2021] provides utilities for converting **BibTeX** entries (via the **R** base function `bibentry()`) to
[CFF]{.underline} files and vice versa. This section describes how the conversion between both formats has been implemented. The crosswalk is
partially based on @Haines_Ruby_CFF_Library_2021[^5].

[^5]: Note that this software performs only the conversion from [CFF]{.underline} to **BibTeX**, however **cffr** can perform the conversion in both
    directions.

In the following two sections, I present an overview of the proposed mapping between the **Entries** and **Fields** of **BibTeX** and the
[CFF]{.underline} keys. After this initial mapping, I propose further transformations to enhance compatibility between both systems using different
[Entry Models](#entrymodels).

For better clarity, when a field is in **bold** (e.g., **\@book, edition**), it corresponds to **BibTeX**, and when the field is
[underlined]{.underline} (e.g., [book, edition]{.underline}), it corresponds to [CFF]{.underline}.

### Entry/Type Crosswalk

For converting general **BibTeX** entries to [CFF]{.underline} types, the following crosswalk is proposed:

```{r etypebibcff, echo=FALSE, results='asis'}
df <- data.frame(
  stringsAsFactors = FALSE,
  a = c(
    "**\\@article**", "**\\@book**",
    "**\\@booklet**", "**\\@conference**", "**\\@inbook**",
    "**\\@incollection**", "**\\@inproceedings**",
    "**\\@manual**", "**\\@mastersthesis**", "**\\@misc**",
    "**\\@phdthesis**", "**\\@proceedings**", "**\\@techreport**",
    "**\\@unpublished**"
  ),
  b = c(
    "[article]{.underline}",
    "[book]{.underline}", "[pamphlet]{.underline}",
    "[conference-paper]{.underline}", "[book]{.underline}",
    "[generic]{.underline}", "[conference-paper]{.underline}",
    "[manual]{.underline}", "[thesis]{.underline}", "[generic]{.underline}",
    "[thesis]{.underline}", "[proceedings]{.underline}",
    "[report]{.underline}", "[unpublished]{.underline}"
  ),
  c = c(
    NA, NA, NA, NA,
    "See [Entry Models](#entrymodels)", "See [Entry Models](#entrymodels)", NA, NA,
    " See [Entry Models](#entrymodels)", NA, " See [Entry Models](#entrymodels)", NA, NA, NA
  )
)

df[is.na(df)] <- ""

knitr::kable(df,
  col.names = c(
    "**BibTeX** Entry", "[**CFF key: type**]{.underline}",
    "Notes"
  ),
  caption = "Entry/Type crosswalk: From **BibTeX** to [CFF]{.underline}"
)
```

Additionally, considering that [CFF]{.underline} allows for a [wide range of
values](https://github.com/citation-file-format/citation-file-format/blob/main/schema-guide.md#definitionsreferencetype) for the [type]{.underline}
field, the following conversion would be applied from [CFF]{.underline} to **BibTeX**:

```{r, echo=FALSE,results='asis'}
df <- data.frame(
  stringsAsFactors = FALSE,
  a = c(
    "[book]{.underline}",
    "[conference, conference-paper]{.underline}",
    "[article, magazine-article, newspaper-article]{.underline}",
    "[manual]{.underline}", "[pamphlet]{.underline}",
    "[proceedings]{.underline}", "[report]{.underline}", "[thesis]{.underline}",
    "[unpublished]{.underline}", "[generic]{.underline}",
    "\\<[any other value]{.underline}\\>"
  ),
  b = c(
    "**\\@book**, **\\@inbook**",
    "**\\@inproceedings**", "**\\@article**",
    "**\\@manual**", "**\\@booklet**", "**\\@proceedings**",
    "**\\@techreport**", "**\\@mastersthesis**,  **\\@phdthesis**",
    "**\\@unpublished**", "**\\@misc**, **\\@incollection**",
    "**\\@misc**"
  ),
  c = c(
    "See [Entry Models](#entrymodels)", NA,
    NA, NA, NA, NA, NA, "See [Entry Models](#entrymodels)", NA,
    "Under specific conditions, see [Entry Models](#entrymodels)", NA
  )
)

df[is.na(df)] <- ""

knitr::kable(df,
  col.names = c(
    "[**CFF key: type**]{.underline}", "**BibTeX** Entry",
    "Notes"
  ),
  caption = "Entry/Type crosswalk: From [CFF]{.underline} to **BibTeX**"
)
```

### Fields/Key Crosswalk

There exists a significant similarity between the definitions and names of certain **BibTeX fields** and [CFF keys]{.underline}. In the following
cases, the equivalence is nearly straightforward:

| **BibTeX Field** | [CFF key]{.underline}                   |
|------------------|-----------------------------------------|
| **address**      | See [Entry Models](#entrymodels)        |
| ***annote***     | \-                                      |
| **author**       | [authors]{.underline}                   |
| **booktitle**    | [collection-title]{.underline}          |
| **chapter**      | [section]{.underline}                   |
| ***crossref***   | \-                                      |
| **edition**      | [edition]{.underline}                   |
| **editor**       | [editors]{.underline}                   |
| **howpublished** | [medium]{.underline}                    |
| **institution**  | See [Entry Models](#entrymodels)        |
| **journal**      | [journal]{.underline}                   |
| ***key***        | \-                                      |
| **month**        | [month]{.underline}                     |
| **note**         | [notes]{.underline}                     |
| **number**       | [issue]{.underline}                     |
| **organization** | See [Entry Models](#entrymodels)        |
| **pages**        | [start]{.underline} & [end]{.underline} |
| **publisher**    | [publisher]{.underline}                 |
| **school**       | See [Entry Models](#entrymodels)        |
| **series**       | See [Entry Models](#entrymodels)        |
| **title**        | [title]{.underline}                     |
| ***type***       | \-                                      |
| **volume**       | [volume]{.underline}                    |
| **year**         | [year]{.underline}                      |

: **BibTeX** - [CFF]{.underline} Field/Key crosswalk

Additionally, there are other [CFF]{.underline} keys that correspond to BibLaTeX fields. We propose to include these fields in the crosswalk[^6], even
though they are not part of the core **BibTeX** fields definition.

[^6]: See @biblatexcheatsheet for a preview of the accepted BibLaTeX fields.

| BibLaTeX Field | [CFF key]{.underline}        |
|----------------|------------------------------|
| **abstract**   | [abstract]{.underline}       |
| **date**       | [date-published]{.underline} |
| **doi**        | [doi]{.underline}            |
| **file**       | [filename]{.underline}       |
| **isbn**       | [isbn]{.underline}           |
| **issn**       | [issn]{.underline}           |
| **issuetitle** | [issue-title]{.underline}    |
| **pagetotal**  | [pages]{.underline}          |
| **translator** | [translators]{.underline}    |
| **url**        | [url]{.underline}            |
| **urldate**    | [date-accessed]{.underline}  |
| **version**    | [version]{.underline}        |

: BibLaTeX - [CFF]{.underline} Field/Key crosswalk

## Entry Models {#entrymodels}

This section presents the specific mapping proposed for each of the \*\*BibTeX entries, providing further information on how each field is treated.
Examples are adapted from the [xampl.bib](https://tug.org/texmf-docs/bibtex/xampl.bib) file provided with the bibtex package [@patashnik].

Note that the **BibTeX** fields marked with "\*" on each table indicate those that are strictly required for each entry type.

### article

The crosswalk of **\@article** does not require any special treatment.

```{r echo=FALSE, results='asis'}
df <- data.frame(
  stringsAsFactors = FALSE,
  a = c(
    "**\\@article**",
    "**author\\***", "**title\\***", "**journal\\***", "**year\\***",
    "**volume**", "**number**", "**pages**", "**month**",
    "**note**"
  ),
  b = c(
    "[type: article, magazine-article,  newspaper-article]{.underline}",
    "[authors]{.underline}", "[title]{.underline}", "[journal]{.underline}",
    "[year]{.underline}", "[volume]{.underline}",
    "[issue]{.underline}", "[start]{.underline} and [end]{.underline}",
    "[month]{.underline}", "[notes]{.underline}"
  ),
  c = c(
    NA, NA, NA, NA, NA, NA, NA,
    "Separated by `--`, i.e, **pages = {3\\--5}** would be  parsed as [start: 3]{.underline}, [end: 5]{.underline}",
    "As a fallback, **month** could be extracted also from  **date** (BibLaTeX field) or  [date-published]{.underline}", NA
  )
)

df[is.na(df)] <- ""


knitr::kable(df,
  col.names = c("**BibTeX**", "[CFF]{.underline}", "Notes"),
  caption = "**\\@article** Model"
)
```


+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| BibTeX                             | [CFF]{.underline}                                | Note                                                     |
+====================================+==================================================+==========================================================+
| **\@article**                      | [type: article, magazine-article,                |                                                          |
|                                    | newspaper-article]{.underline}                   |                                                          |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **author\***                       | [authors]{.underline}                            |                                                          |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **title\***                        | [title]{.underline}                              |                                                          |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **journal\***                      | [journal]{.underline}                            |                                                          |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **year\***                         | [year]{.underline}                               |                                                          |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **volume**                         | [volume]{.underline}                             |                                                          |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **number**                         | [issue]{.underline}                              |                                                          |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **pages**                          | [start]{.underline} and [end]{.underline}        | Separated by `--`, i.e, **pages = {3\--5}** would be     |
|                                    |                                                  | parsed as:                                               |
|                                    |                                                  |                                                          |
|                                    |                                                  | -   [start: 3]{.underline}                               |
|                                    |                                                  |                                                          |
|                                    |                                                  | -   [end: 5]{.underline}                                 |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **month**                          | [month]{.underline}                              | As a fallback, **month** could be extracted also from    |
|                                    |                                                  | **date** (BibLaTeX field) or                             |
|                                    |                                                  | [date-published]{.underline}                             |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **note**                           | [notes]{.underline}                              |                                                          |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+

: **\@article** Model

**Examples**

[*BibTeX entry*]{.underline}

``` bibtex
@article{article-full,
    title        = {The Gnats and Gnus Document Preparation System},
    author       = {Leslie A. Aamport},
    year         = 1986,
    month        = jul,
    journal      = {{G-Animal's} Journal},
    volume       = 41,
    number       = 7,
    pages        = {73+},
    note         = {This is a full ARTICLE entry}
}
```

[*CFF entry*]{.underline}

```{r echo=FALSE}
bib <- "@article{article-full,
    title        = {The Gnats and Gnus Document Preparation System},
    author       = {Leslie A. Aamport},
    year         = 1986,
    month        = jul,
    journal      = {{G-Animal's} Journal},
    volume       = 41,
    number       = 7,
    pages        = {73+},
    note         = {This is a full ARTICLE entry}}"

cff_from_bibtex(bib)
```

[*From CFF to BibTeX*]{.underline}

```{r echo=FALSE}
toBibtex(cff_to_bibtex(cff_from_bibtex(bib)))
```

### book/inbook

In terms of field required on BibTeX, the only difference between **\@book** and **\@inbook** is that the latter requires also a **chapter** or
**pages**, while for **\@book** these fields are not even optional. So we propose here to identify an **\@inbook** on CFF as a [book]{.underline} with
[section]{.underline} and [start]{.underline}-[end]{.underline} fields (CFF).

Another specificity is that **series** field is mapped to [collection-title]{.underline} and **address** is mapped as the [address]{.underline} of the
[publisher]{.underline} (CFF).

+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| BibTeX                            | [CFF]{.underline}                 | Note                                                                     |
+===================================+===================================+==========================================================================+
| **\@book**                        | [type: book]{.underline}          |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **\@inbook**                      | [type: book]{.underline}          | For identifying an **\@inbook** in CFF, assess if [section]{.underline}  |
|                                   |                                   | or [start/end]{.underline} information is available                      |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **author\***                      | [authors]{.underline}             |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **editor\***                      | [editors]{.underline}             |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **title\***                       | [title]{.underline}               |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **publisher\***                   | [publisher]{.underline}           |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **year\***                        | [year]{.underline}                |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **chapter\***                     | [section]{.underline}             | Only required on **\@inbook**                                            |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **pages\***                       | [start]{.underline} and           | Only required on **\@inbook**                                            |
|                                   | [end]{.underline}                 |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **volume**                        | [volume]{.underline}              |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **number**                        | [issue]{.underline}               |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **series**                        | [coll ection-title]{.underline}   |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **address**                       | [address]{.underline} property of | As a fallback, the field [location]{.underline} can be used              |
|                                   | [publisher]{.underline}           |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **edition**                       | [edition]{.underline}             |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **month**                         | [month]{.underline}               | See **Note** on [article]                                                |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **note**                          | [notes]{.underline}               |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+

: **\@book/\@inbook** Model

**Examples: book**

[*BibTeX entry*]{.underline}

``` bibtex
@book{book-full,
    title        = {Seminumerical Algorithms},
    author       = {Donald E. Knuth},
    year         = 1981,
    month        = 10,
    publisher    = {Addison-Wesley},
    address      = {Reading, Massachusetts},
    series       = {The Art of Computer Programming},
    volume       = 2,
    note         = {This is a full BOOK entry},
    edition      = {Second}
}
```

[*CFF entry*]{.underline}

```{r echo=FALSE}
bib <- "@book{book-full,
    title        = {Seminumerical Algorithms},
    author       = {Donald E. Knuth},
    year         = 1981,
    month        = 10,
    publisher    = {Addison-Wesley},
    address      = {Reading, Massachusetts},
    series       = {The Art of Computer Programming},
    volume       = 2,
    note         = {This is a full BOOK entry},
    edition      = {Second}
}"

cff_from_bibtex(bib)
```

[*From CFF to BibTeX*]{.underline}

```{r,  echo=FALSE}
toBibtex(cff_to_bibtex(cff_from_bibtex(bib)))
```

**Examples: inbook**

[*BibTeX entry*]{.underline}

``` bibtex
@inbook{inbook-full,
    title        = {Fundamental Algorithms},
    author       = {Donald E. Knuth},
    year         = 1973,
    month        = 10,
    publisher    = {Addison-Wesley},
    address      = {Reading, Massachusetts},
    series       = {The Art of Computer Programming},
    volume       = 1,
    pages        = {10--119},
    note         = {This is a full INBOOK entry},
    edition      = {Second},
    type         = {Section},
    chapter      = {1.2}
}
```

[*CFF entry*]{.underline}

```{r echo=FALSE,}
bib <- "@inbook{inbook-full,
    title        = {Fundamental Algorithms},
    author       = {Donald E. Knuth},
    year         = 1973,
    month        = 10,
    publisher    = {Addison-Wesley},
    address      = {Reading, Massachusetts},
    series       = {The Art of Computer Programming},
    volume       = 1,
    pages        = {10--119},
    note         = {This is a full INBOOK entry},
    edition      = {Second},
    type         = {Section},
    chapter      = {1.2}
}"

cff_from_bibtex(bib)
```

[*From CFF to BibTeX*]{.underline}

```{r echo=FALSE,}
toBibtex(cff_to_bibtex(cff_from_bibtex(bib)))
```

### booklet

In **\@booklet** **address** is mapped to [location]{.underline}.

+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| BibTeX                            | [CFF]{.underline}                 | Note                                                                     |
+===================================+===================================+==========================================================================+
| **\@booklet**                     | [type: pa mphlet]{.underline}     |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **title\***                       | [title]{.underline}               |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **author\***                      | [a uthors]{.underline}            |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **howpublished**                  | [medium]{.underline}              |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **address**                       | [lo cation]{.underline}           |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **month**                         | [month]{.underline}               | See **Note** on [article]                                                |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **year**                          | [year]{.underline}                | As a fallback, **year** could be extracted also from **date** (BibLaTeX  |
|                                   |                                   | field)/ [date-published]{.underline}                                     |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **note**                          | [notes]{.underline}               |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+

: **\@booklet** Model

**Examples**

[*BibTeX entry*]{.underline}

``` bibtex
@booklet{booklet-full,
    title        = {The Programming of Computer Art},
    author       = {Jill C. Knvth},
    date         = {1988-03-14},
    month        = feb,
    address      = {Stanford, California},
    note         = {This is a full BOOKLET entry},
    howpublished = {Vernier Art Center}
}
```

[*CFF entry*]{.underline}

```{r echo=FALSE, }
bib <- "@booklet{booklet-full,
    title        = {The Programming of Computer Art},
    author       = {Jill C. Knvth},
    date         = {1988-03-14},
    month        = feb,
    address      = {Stanford, California},
    note         = {This is a full BOOKLET entry},
    howpublished = {Vernier Art Center}
}"

cff_from_bibtex(bib)
```

[*From CFF to BibTeX*]{.underline}

```{r echo=FALSE, }
toBibtex(cff_to_bibtex(cff_from_bibtex(bib)))
```

### conference/inproceedings

Note that in this case, **organization** is mapped to [institution]{.underline}, as BibTeX does not prescribe the use of **institution** on these
entries.

+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| BibTeX                            | CFF                               | Note                                                                     |
+===================================+===================================+==========================================================================+
| **\@conference /                  | [type: con                        | CFF entries with [type; conference]{.underline} are mapped back to       |
| \@inproceedings**                 | ference-paper]{.underline}        | **\@inproceedings**.                                                     |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **author\***                      | [authors]{.underline}             |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **title\***                       | [title]{.underline}               |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **booktitle\***                   | [col lection-title]{.underline}   |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **year\***                        | [year]{.underline}                |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **editor**                        | [editors]{.underline}             |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **volume**                        | [volume]{.underline}              |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **number**                        | [issue]{.underline}               |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **series**                        | [conference]{.underline}          |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **pages**                         | [start]{.underline} and           | See **Note** on [article]                                                |
|                                   | [end]{.underline}                 |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **address**                       | [location]{.underline}            | As a fallback, [address]{.underline} property of                         |
|                                   |                                   | [conference]{.underline} can be used                                     |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **month**                         | [month]{.underline}               | See **Note** on [article]                                                |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **organization**                  | [institution]{.underline}         |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **publisher**                     | [publisher]{.underline}           |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **note**                          | [notes]{.underline}               |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+

: **\@conference/\@inproceedings** Model

**Examples**

[*BibTeX entry*]{.underline}

``` bibtex
@inproceedings{inproceedings-full,
    title        = {On Notions of Information Transfer in {VLSI} Circuits},
    author       = {Alfred V. Oaho and Jeffrey D. Ullman and Mihalis Yannakakis},
    year         = 1983,
    month        = mar,
    booktitle    = {Proc. Fifteenth Annual ACM Symposium on the Theory of Computing},
    publisher    = {Academic Press},
    address      = {Boston},
    series       = {All ACM Conferences},
    number       = 17,
    pages        = {133--139},
    editor       = {Wizard V. Oz and Mihalis Yannakakis},
    organization = {The OX Association for Computing Machinery}
}
```

[*CFF entry*]{.underline}

```{r echo=FALSE,}
bib <- "@inproceedings{inproceedings-full,
    title        = {On Notions of Information Transfer in {VLSI} Circuits},
    author       = {Alfred V. Oaho and Jeffrey D. Ullman and Mihalis Yannakakis},
    year         = 1983,
    month        = mar,
    booktitle    = {Proc. Fifteenth Annual ACM Symposium on the Theory of Computing},
    publisher    = {Academic Press},
    address      = {Boston},
    series       = {All ACM Conferences},
    number       = 17,
    pages        = {133--139},
    editor       = {Wizard V. Oz and Mihalis Yannakakis},
    organization = {The OX Association for Computing Machinery}
}"

cff_from_bibtex(bib)
```

[*From CFF to BibTeX*]{.underline}

```{r echo=FALSE,}
toBibtex(cff_to_bibtex(cff_from_bibtex(bib)))
```

### incollection

As **booktitle** is a required field, we propose to map that field to [collection-title]{.underline} and the [type]{.underline} to
[generic]{.underline}. Therefore, an **\@incollection** is a [type: generic]{.underline} with a [collection-title]{.underline} key.

+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **BibTeX**                         | [CFF]{.underline}                                          | Note                                            |
+====================================+============================================================+=================================================+
| **\@incollection**                 | [type: generic]{.underline}                                | Including a [collection-title]{.underline}      |
|                                    |                                                            | value                                           |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **author\***                       | [authors]{.underline}                                      |                                                 |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **title\***                        | [title]{.underline}                                        |                                                 |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **booktitle\***                    | [collection-title]{.underline}                             |                                                 |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **publisher\***                    | [publisher]{.underline}                                    |                                                 |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **year\***                         | [year]{.underline}                                         |                                                 |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **editor**                         | [editors]{.underline}                                      |                                                 |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **volume**                         | [volume]{.underline}                                       |                                                 |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **number**                         | [issue]{.underline}                                        |                                                 |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **series**                         | [series]{.underline}                                       |                                                 |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **type**                           | \-                                                         | Ignored                                         |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **chapter**                        | [section]{.underline}                                      |                                                 |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **pages**                          | [start]{.underline} and [end]{.underline}                  | See **Note** on [article]                       |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **address**                        | [address]{.underline} property of [publisher]{.underline}  | See **Note** on [book/inbook]                   |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **edition**                        | [edition]{.underline}                                      |                                                 |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **month**                          | [month]{.underline}                                        | See **Note** on [article]                       |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **note**                           | [notes]{.underline}                                        |                                                 |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+

: **\@incollection** Model

**Examples**

[*BibTeX entry*]{.underline}

``` bibtex
@incollection{incollection-full,
    title        = {Semigroups of Recurrences},
    author       = {Daniel D. Lincoll},
    year         = 1977,
    month        = sep,
    booktitle    = {High Speed Computer and Algorithm Organization},
    publisher    = {Academic Press},
    address      = {New York},
    series       = {Fast Computers},
    number       = 23,
    pages        = {179--183},
    note         = {This is a full INCOLLECTION entry},
    editor       = {David J. Lipcoll and D. H. Lawrie and A. H. Sameh},
    chapter      = 3,
    type         = {Part},
    edition      = {Third}
}
```

[*CFF entry*]{.underline}

```{r echo=FALSE,}
bib <- "@incollection{incollection-full,
    title        = {Semigroups of Recurrences},
    author       = {Daniel D. Lincoll},
    year         = 1977,
    month        = sep,
    booktitle    = {High Speed Computer and Algorithm Organization},
    publisher    = {Academic Press},
    address      = {New York},
    series       = {Fast Computers},
    number       = 23,
    pages        = {179--183},
    note         = {This is a full INCOLLECTION entry},
    editor       = {David J. Lipcoll and D. H. Lawrie and A. H. Sameh},
    chapter      = 3,
    type         = {Part},
    edition      = {Third}
}"

cff_from_bibtex(bib)
```

[*From CFF to BibTeX*]{.underline}

```{r echo=FALSE,}
toBibtex(cff_to_bibtex(cff_from_bibtex(bib)))
```

### manual

As in the case of [conference/inproceedings], **organization** is mapped to [institution]{.underline}.

+------------------------------------+-------------------------------------------------------+------------------------------------------------------+
| **BibTeX**                         | [CFF]{.underline}                                     | Note                                                 |
+====================================+=======================================================+======================================================+
| **\@manual**                       | [type: manual]{.underline}                            |                                                      |
+------------------------------------+-------------------------------------------------------+------------------------------------------------------+
| **title\***                        | [title]{.underline}                                   |                                                      |
+------------------------------------+-------------------------------------------------------+------------------------------------------------------+
| **author**                         | [authors]{.underline}                                 |                                                      |
+------------------------------------+-------------------------------------------------------+------------------------------------------------------+
| **organization**                   | [institution]{.underline}                             |                                                      |
+------------------------------------+-------------------------------------------------------+------------------------------------------------------+
| **address**                        | [location]{.underline}                                |                                                      |
+------------------------------------+-------------------------------------------------------+------------------------------------------------------+
| **edition**                        | [edition]{.underline}                                 |                                                      |
+------------------------------------+-------------------------------------------------------+------------------------------------------------------+
| **month**                          | [month]{.underline}                                   | See **Note** on [article]                            |
+------------------------------------+-------------------------------------------------------+------------------------------------------------------+
| **year**                           | [year]{.underline}                                    | See **Note** on [booklet]                            |
+------------------------------------+-------------------------------------------------------+------------------------------------------------------+
| **note**                           | [notes]{.underline}                                   |                                                      |
+------------------------------------+-------------------------------------------------------+------------------------------------------------------+

: **\@manual** Model

**Examples**

[*BibTeX entry*]{.underline}

Note that **month** can't be parsed to a single integer in the range `1--12` as required on CFF, so it is not parsed to avoid validation errors.

``` bibtex
@manual{manual-full,
  title        = {The Definitive Computer Manual},
    author       = {Larry Manmaker},
    year         = 1986,
    month        = {apr-may},
    address      = {Silicon Valley},
    note         = {This is a full MANUAL entry},
    organization = {Chips-R-Us},
    edition      = {Silver}
}
```

[*CFF entry*]{.underline}

```{r echo=FALSE,}
bib <- "@manual{manual-full,
  title        = {The Definitive Computer Manual},
    author       = {Larry Manmaker},
    year         = 1986,
    month        = {apr-may},
    address      = {Silicon Valley},
    note         = {This is a full MANUAL entry},
    organization = {Chips-R-Us},
    edition      = {Silver}
}"

cff_from_bibtex(bib)
```

[*From CFF to BibTeX*]{.underline}

```{r echo=FALSE,}
toBibtex(cff_to_bibtex(cff_from_bibtex(bib)))
```

### mastersthesis/phdthesis

In terms of field required on BibTeX, it is identical for both **\@mastersthesis** and **\@phdthesis.**

We propose here to identify each type of thesis using the field [thesis-type]{.underline} (CFF). So if [thesis-type]{.underline} contains a [regex
pattern](https://regex101.com/r/mBWfbs/1) `(?i)(phd)` it would be recognized as **\@phdthesis**.

+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| BibTeX                             | [CFF]{.underline}                                | Note                                                     |
+====================================+==================================================+==========================================================+
| **\@mastersthesis**                | [type: thesis]{.underline}                       | Use also [thesis-type]{.underline} for identifying the   |
|                                    |                                                  | thesis type.                                             |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **\@phdthesis**                    | [type: thesis]{.underline}                       | Use also [thesis-type]{.underline} for identifying the   |
|                                    |                                                  | thesis type.                                             |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **author\***                       | [authors]{.underline}                            |                                                          |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **title\***                        | [title]{.underline}                              |                                                          |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **school\***                       | [institution]{.underline}                        |                                                          |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **year\***                         | [year]{.underline}                               |                                                          |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **type**                           |                                                  |                                                          |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **address**                        | [address]{.underline} property of                | See **Note** on [book/inbook]                            |
|                                    | [institution]{.underline}                        |                                                          |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **month**                          | [month]{.underline}                              | See **Note** on [article]                                |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+
| **note**                           | [notes]{.underline}                              |                                                          |
+------------------------------------+--------------------------------------------------+----------------------------------------------------------+

: **\@mastersthesis/phdthesis** Model

**Examples: mastersthesis**

[*BibTeX entry*]{.underline}

``` bibtex
@mastersthesis{mastersthesis-full,
    title        = {Mastering Thesis Writing},
    author       = {Edouard Masterly},
    year         = 1988,
    month        = jun,
    address      = {English Department},
    note         = {This is a full MASTERSTHESIS entry},
    school       = {Stanford University},
    type         = {Master's project}
}
```

[*CFF entry*]{.underline}

```{r echo=FALSE}
bib <- "@mastersthesis{mastersthesis-full,
    title        = {Mastering Thesis Writing},
    author       = {Edouard Masterly},
    year         = 1988,
    month        = jun,
    address      = {English Department},
    note         = {This is a full MASTERSTHESIS entry},
    school       = {Stanford University},
    type         = {Master's project}
}"

cff_from_bibtex(bib)
```

[*From CFF to BibTeX*]{.underline}

```{r,  echo=FALSE}
toBibtex(cff_to_bibtex(cff_from_bibtex(bib)))
```

**Examples: phdthesis**

[*BibTeX entry*]{.underline}

``` bibtex
@phdthesis{phdthesis-full,
    title        = {Fighting Fire with Fire: Festooning {F}rench Phrases},
    author       = {F. Phidias Phony-Baloney},
    year         = 1988,
    month        = jun,
    address      = {Department of French},
    note         = {This is a full PHDTHESIS entry},
    school       = {Fanstord University},
    type         = {{PhD} Dissertation}
}
```

[*CFF entry*]{.underline}

```{r echo=FALSE,}
bib <- "@phdthesis{phdthesis-full,
    title        = {Fighting Fire with Fire: Festooning {F}rench Phrases},
    author       = {F. Phidias Phony-Baloney},
    year         = 1988,
    month        = jun,
    address      = {Department of French},
    note         = {This is a full PHDTHESIS entry},
    school       = {Fanstord University},
    type         = {{PhD} Dissertation}
}"

cff_from_bibtex(bib)
```

[*From CFF to BibTeX*]{.underline}

```{r echo=FALSE,}
toBibtex(cff_to_bibtex(cff_from_bibtex(bib)))
```

### misc

The crosswalk of **\@misc** does not require any special treatment. This entry does not require any field.

Note also that it is mapped to [type: generic]{.underline} as [incollection], but in this case **booktitle** is not even an option, so the proposed
definition should cover both **\@misc** and **\@incollection** without problems.

+------------------------------------+--------------------------------------------------------+----------------------------------------------------+
| **BibTeX**                         | [CFF]{.underline}                                      | Note                                               |
+====================================+========================================================+====================================================+
| **\@misc**                         | [type: generic]{.underline}                            |                                                    |
+------------------------------------+--------------------------------------------------------+----------------------------------------------------+
| **author**                         | [authors]{.underline}                                  |                                                    |
+------------------------------------+--------------------------------------------------------+----------------------------------------------------+
| **title**                          | [title]{.underline}                                    |                                                    |
+------------------------------------+--------------------------------------------------------+----------------------------------------------------+
| **howpublished**                   | [medium]{.underline}                                   |                                                    |
+------------------------------------+--------------------------------------------------------+----------------------------------------------------+
| **month**                          | [month]{.underline}                                    | See **Note** on [article]                          |
+------------------------------------+--------------------------------------------------------+----------------------------------------------------+
| **year**                           | [year]{.underline}                                     | See **Note** on [booklet]                          |
+------------------------------------+--------------------------------------------------------+----------------------------------------------------+
| **note**                           | [notes]{.underline}                                    |                                                    |
+------------------------------------+--------------------------------------------------------+----------------------------------------------------+

: **\@misc** Model

**Examples**

[*BibTeX entry*]{.underline}

``` bibtex
@misc{misc-full,
    title        = {Handing out random pamphlets in airports},
    author       = {Joe-Bob Missilany},
    year         = 1984,
    month        = oct,
    note         = {This is a full MISC entry},
    howpublished = {Handed out at O'Hare}
}
```

[*CFF entry*]{.underline}

```{r echo=FALSE,}
bib <- "@misc{misc-full,
    title        = {Handing out random pamphlets in airports},
    author       = {Joe-Bob Missilany},
    year         = 1984,
    month        = oct,
    note         = {This is a full MISC entry},
    howpublished = {Handed out at O'Hare}
}"

cff_from_bibtex(bib)
```

[*From CFF to BibTeX*]{.underline}

```{r echo=FALSE,}
toBibtex(cff_to_bibtex(cff_from_bibtex(bib)))
```

### proceedings

The proposed model is similar to [conference/inproceedings]. Note that **\@proceedings** does not prescribe a **author** field. On this cases, as
[authors]{.underline} is required on CFF, we would use *anonymous*[^7] when converting to CFF and omit it on the conversion back to CFF.

[^7]: As proposed on [*How to deal with unknown individual
    authors?*](https://github.com/citation-file-format/citation-file-format/blob/main/schema-guide.md#how-to-deal-with-unknown-individual-authors),
    **(Guide to Citation File Format schema version 1.2.0)**

+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| BibTeX                            | [CFF]{.underline}                 | Note                                                                     |
+===================================+===================================+==========================================================================+
| **\@proceedings**                 | [type: pro ceedings]{.underline}  |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **title\***                       | [title]{.underline}               |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **year\***                        | [year]{.underline}                |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **editor**                        | [editors]{.underline}             |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **volume**                        | [volume]{.underline}              |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **number**                        | [issue]{.underline}               |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **series**                        | [co nference]{.underline}         |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **address**                       | [location]{.underline}            | As a fallback, [address]{.underline} property of                         |
|                                   |                                   | [conference]{.underline} can be used                                     |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **month**                         | [month]{.underline}               | See **Note** on [article]                                                |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **organization**                  | [ins titution]{.underline}        |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **publisher**                     | [p ublisher]{.underline}          |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+
| **note**                          | [notes]{.underline}               |                                                                          |
+-----------------------------------+-----------------------------------+--------------------------------------------------------------------------+

: **\@conference/\@inproceedings** Model

**Examples**

[*BibTeX entry*]{.underline}

``` bibtex
@proceedings{proceedings-full,
    title        = {Proc. Fifteenth Annual ACM Symposium on the Theory of Computing},
    year         = 1983,
    month        = mar,
    publisher    = {Academic Press},
    address      = {Boston},
    series       = {All ACM Conferences},
    number       = 17,
    note         = {This is a full PROCEEDINGS entry},
    editor       = {Wizard V. Oz and Mihalis Yannakakis},
    organization = {The OX Association for Computing Machinery}
}
```

[*CFF entry*]{.underline}

```{r echo=FALSE,}
bib <- "@proceedings{proceedings-full,
    title        = {Proc. Fifteenth Annual ACM Symposium on the Theory of Computing},
    year         = 1983,
    month        = mar,
    publisher    = {Academic Press},
    address      = {Boston},
    series       = {All ACM Conferences},
    number       = 17,
    note         = {This is a full PROCEEDINGS entry},
    editor       = {Wizard V. Oz and Mihalis Yannakakis},
    organization = {The OX Association for Computing Machinery}
}"

cff_from_bibtex(bib)
```

[*From CFF to BibTeX*]{.underline}

```{r echo=FALSE,}
toBibtex(cff_to_bibtex(cff_from_bibtex(bib)))
```

### techreport

+-----------------------------------+--------------------------------------------------------------------------+-----------------------------------+
| **BibTeX**                        | [CFF]{.underline}                                                        | Note                              |
+===================================+==========================================================================+===================================+
| **\@techreport**                  | [type: report]{.underline}                                               |                                   |
+-----------------------------------+--------------------------------------------------------------------------+-----------------------------------+
| **author\***                      | [authors]{.underline}                                                    |                                   |
+-----------------------------------+--------------------------------------------------------------------------+-----------------------------------+
| **title\***                       | [title]{.underline}                                                      |                                   |
+-----------------------------------+--------------------------------------------------------------------------+-----------------------------------+
| **institution\***                 | [institution]{.underline}                                                |                                   |
+-----------------------------------+--------------------------------------------------------------------------+-----------------------------------+
| **year\***                        | [year]{.underline}                                                       |                                   |
+-----------------------------------+--------------------------------------------------------------------------+-----------------------------------+
| **type**                          | \-                                                                       | Ignored                           |
+-----------------------------------+--------------------------------------------------------------------------+-----------------------------------+
| **number**                        | [issue]{.underline}                                                      |                                   |
+-----------------------------------+--------------------------------------------------------------------------+-----------------------------------+
| **address**                       | [address]{.underline} property of [institution]{.underline}              | See **Note** on [book/inbook]     |
+-----------------------------------+--------------------------------------------------------------------------+-----------------------------------+
| **month**                         | [month]{.underline}                                                      | See **Note** on [article]         |
+-----------------------------------+--------------------------------------------------------------------------+-----------------------------------+
| **note**                          | [notes]{.underline}                                                      |                                   |
+-----------------------------------+--------------------------------------------------------------------------+-----------------------------------+

: **\@techreport** Model

**Examples**

[*BibTeX entry*]{.underline}

``` bibtex
@techreport{techreport-full,
    title        = {A Sorting Algorithm},
    author       = {Tom Terrific},
    year         = 1988,
    month        = oct,
    address      = {Computer Science Department, Fanstord, California},
    number       = 7,
    note         = {This is a full TECHREPORT entry},
    institution  = {Fanstord University},
    type         = {Wishful Research Result}
}
```

[*CFF entry*]{.underline}

```{r echo=FALSE,}
bib <- "@techreport{techreport-full,
    title        = {A Sorting Algorithm},
    author       = {Tom Terrific},
    year         = 1988,
    month        = oct,
    address      = {Computer Science Department, Fanstord, California},
    number       = 7,
    note         = {This is a full TECHREPORT entry},
    institution  = {Fanstord University},
    type         = {Wishful Research Result}
}"

cff_from_bibtex(bib)
```

[*From CFF to BibTeX*]{.underline}

```{r echo=FALSE,}
toBibtex(cff_to_bibtex(cff_from_bibtex(bib)))
```

### unpublished

+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **BibTeX**                         | [CFF]{.underline}                                          | Note                                            |
+====================================+============================================================+=================================================+
| **\@unpublished**                  | [type: unpublished]{.underline}                            |                                                 |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **author\***                       | [authors]{.underline}                                      |                                                 |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **title\***                        | [title]{.underline}                                        |                                                 |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **note\***                         | [notes]{.underline}                                        |                                                 |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **month**                          | [month]{.underline}                                        | See **Note** on [article]                       |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+
| **year**                           | [year]{.underline}                                         | See **Note** on [booklet]                       |
+------------------------------------+------------------------------------------------------------+-------------------------------------------------+

: **\@unpublished** Model

**Examples**

[*BibTeX entry*]{.underline}

``` bibtex
@unpublished{unpublished-minimal,
    title        = {Lower Bounds for Wishful Research Results},
    author       = {Ulrich Underwood and Ned Net and Paul Pot},
    note         = {Talk at Fanstord University (this is a minimal UNPUBLISHED entry)}
}
```

[*CFF entry*]{.underline}

```{r echo=FALSE,}
bib <- "@unpublished{unpublished-minimal,
    title        = {Lower Bounds for Wishful Research Results},
    author       = {Ulrich Underwood and Ned Net and Paul Pot},
    note         = {Talk at Fanstord University (this is a minimal UNPUBLISHED entry)}
}"

cff_from_bibtex(bib)
```

[*From CFF to BibTeX*]{.underline}

```{r echo=FALSE,}
toBibtex(cff_to_bibtex(cff_from_bibtex(bib)))
```

## References
