---
title: "Crosswalk"
subtitle: "From R to CFF"
author: Diego Hernangómez
bibliography: bibtex.bib
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Crosswalk}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(cffr)

```

The goal of this vignette is to provide clarity on how **cffr** extracts the
metadata of a package and parse it to a valid `CITATION.cff` file.


## Summary

we summarize here the fields that **cffr** can parse and the original source of
information for each one of them. The details on each key are presented on the
next section of the document. The assessment of fields are based om the [Guide
to Citation File Format schema version
1.2.0](https://github.com/citation-file-format/citation-file-format/blob/main/schema-guide.md#valid-keys) [@Druskat_Citation_File_Format_2021].

```{r summary , echo=FALSE}

keys <- cff_schema_keys(sorted = TRUE)
origin <- vector(length = length(keys))
origin[keys == "cff-version"] <- "parameter on function"
origin[keys == "type"] <- "Fixed value: 'software'"
origin[keys == "identifiers"] <- "DESCRIPTION/CITATION files"

origin[keys %in% c(
  "message",
  "title",
  "version",
  "authors",
  "abstract",
  "repository",
  "repository-code",
  "url",
  "date-released",
  "contact",
  "keywords",
  "license"
)] <- "DESCRIPTION file"

origin[keys %in% c(
  "doi",
  "preferred-citation",
  "references"
)] <- "CITATION file"


origin[origin == FALSE] <- "Not parsed by cffr"

df <- data.frame(
  key = paste0("<a href='#", keys, "'>", keys, "</strong>"),
  source = origin
)


knitr::kable(df, escape = FALSE)
```

## Details

### abstract

This key is extracted from the "Description" field of the DESCRIPTION file.

<details><summary>Example</summary>


```{r abstract}

library(cffr)

# Create cffr for yaml

cff_obj <- cff_create("rmarkdown")

# Get DESCRIPTION of rmarkdown to check

pkg <- desc::desc(file.path(find.package("rmarkdown"), "DESCRIPTION"))

cat(cff_obj$abstract)

cat(pkg$get("Description"))
```

</details>

### authors

This key is parsed from the "Authors" or "Authors\@R" field of the DESCRIPTION
file. Only persons with the role "aut" or "cre" are considered.

<details><summary>Example</summary>

```{r authors}

cff_obj <- cff_create("rmarkdown")
pkg <- desc::desc(file.path(find.package("rmarkdown"), "DESCRIPTION"))

cff_obj$authors

authors <- pkg$get_authors()

authors[vapply(authors, function(x) {
  "aut" %in% x$role || "cre" %in% x$role
}, logical(1))]
```

</details>

### cff-version

This value can be set via the parameters of the `cff_create()`/`cff_write()`
functions:

<details><summary>Example</summary>

```{r cffversion}

cff_objv110 <- cff_create("jsonlite", cff_version = "v1.1.0")

cat(cff_objv110$`cff-version`)
```

</details>

### commit

This value is not extracted from the metadata of the package. See the
description on the [Guide
to CFF schema v1.2.0](https://github.com/citation-file-format/citation-file-format/blob/main/schema-guide.md#commit).

> - **description**: The commit hash or revision number of the software version.
> - **usage**:<br><br>
>
>     ```yaml
>     commit: 1ff847d81f29c45a3a1a5ce73d38e45c2f319bba
>
>     commit: "Revision: 8612"
>     ```
>


### contact

This key is parsed from the "Authors" or "Authors\@R" field of the DESCRIPTION
file. Only persons with the role "cre" (i.e, the maintainer(s)) are considered.

<details><summary>Example</summary>

```{r contact}

cff_obj <- cff_create("rmarkdown")
pkg <- desc::desc(file.path(find.package("rmarkdown"), "DESCRIPTION"))

cff_obj$contact

authors[vapply(authors, function(x) {
  "cre" %in% x$role
}, logical(1))]
```
</details>

### date-released

This key is parsed from the "Date" field or, if not present, from the 
"Date/Publication" field that is present on packages built on CRAN.


<details><summary>Example</summary>


```{r date-released}

# From an installed package

cff_obj <- cff_create("rmarkdown")
pkg <- desc::desc(file.path(find.package("rmarkdown"), "DESCRIPTION"))

pkg$get("Date/Publication")
cff_obj$`date-released`

# A DESCRIPTION file without a Date
nodate <- system.file("examples/DESCRIPTION_basic", package = "cffr")
tmp <- tempfile("DESCRIPTION")

# Create a temporary file
file.copy(nodate, tmp)


pkgnodate <- desc::desc(tmp)
cffnodate <- cff_create(tmp)

# Won't appear
cffnodate$`date-released`

pkgnodate

# Adding a Date

desc::desc_set("Date", "1999-01-01", file = tmp)

cff_create(tmp)$`date-released`
```
</details>


### doi

This key is parsed from the "doi" field of the 
[preferred-citation](#preferred-citation) object.

<details><summary>Example</summary>

```{r doi}

cff_doi <- cff_create("cffr")

cff_doi$doi

cff_doi$`preferred-citation`$doi
```

</details>


### identifiers

This key includes all the possible identifiers of the package:
-  From the DESCRIPTION field, it includes all the urls not included in
   [url](#url) or [repository-code](#repository-code).
- From the CITATION file, it includes all the dois not included in [doi](#doi)
and the identifiers (if any) not included in the "identifiers" key of 
[preferred-citation](#preferred-citation).

<details><summary>Example</summary>

```{r identifiers}
file <- system.file("examples/DESCRIPTION_many_urls", package = "cffr")

pkg <- desc::desc(file)
pkg$get_urls()

cat(cff_create(file)$url)
cat(cff_create(file)$`repository-code`)
cff_create(file)$identifiers
```
</details>

### keywords

This key is extracted from the DESCRIPTION file. The keywords should appear
in the DESCRIPTION as:
```
...
X-schema.org-keywords: keyword1, keyword2, keyword3

```

<details><summary>Example</summary>

```{r keyword}

# A DESCRIPTION file without keywords
nokeywords <- system.file("examples/DESCRIPTION_basic", package = "cffr")
tmp2 <- tempfile("DESCRIPTION")

# Create a temporary file
file.copy(nokeywords, tmp2)


pkgnokeywords <- desc::desc(tmp2)
cffnokeywords <- cff_create(tmp2)

# Won't appear
cffnokeywords$keywords

pkgnokeywords

# Adding Keywords

desc::desc_set("X-schema.org-keywords", "keyword1, keyword2, keyword3", file = tmp2)

cff_create(tmp2)$keywords
```


</details>


### license

This key is extracted from the "License" field of the DESCRIPTION file.

<details><summary>Example</summary>


```{r license}

cff_obj <- cff_create("yaml")

cff_obj$license

pkg <- desc::desc(file.path(find.package("yaml"), "DESCRIPTION"))

pkg$get("License")
```


</details>

### license-url

This key is not extracted from the metadata of the package. See the
description on the [Guide
to CFF schema v1.2.0](https://github.com/citation-file-format/citation-file-format/blob/main/schema-guide.md#license-url).


> - **description**: The URL of the license text under which the software or 
> dataset is licensed (only for non-standard licenses not included in the 
> [SPDX License List](https://github.com/citation-file-format/citation-file-format/blob/main/schema-guide.md#definitionslicense-enum)).
> - **usage**:<br><br>
>     ```yaml
>     license-url: "https://obscure-licenses.com?id=1234"
>     ```


### message

This key is extracted from the DESCRIPTION field, specifically as:

```{r eval=FALSE}
msg <- paste0(
  'To cite package "',
  "NAME_OF_THE_PACKAGE",
  '" in publications use:'
)
```

<details><summary>Example</summary>


```{r message}

cff_create("jsonlite")$message
```
</details>

### preferred-citation

This key is extracted from the CITATION file. If several references are 
provided, it would select the first citation as the "preferred-citation" and
the rest of them as [references](#references).


<details><summary>Example</summary>


```{r preferred-citation}

cffobj <- cff_create("rmarkdown")

cffobj$`preferred-citation`

citation("rmarkdown")[1]
```
</details>


### references

This key is extracted from the CITATION file if several references are 
provided. The first citation is considered as the 
[preferred-citation](#preferred-citation) and the rest of them as "references".


<details><summary>Example</summary>


```{r references}

cffobj <- cff_create("rmarkdown")

cffobj$references

print(citation("rmarkdown")[-1], bibtex = TRUE)
```
</details>


### repository

This key is extracted from the "Repository" field of the DESCRIPTION file. 
Usually, this field is auto-populated when a package is hosted on a repo (like
CRAN or the [r-universe](https://r-universe.dev/). For packages without this
field on the DESCRIPTION (that is the typical case for an in-development
package), **cffr** would try to search the package on any of the 
default repositories specified on  `options("repos")`.

If **cffr** detects that the package is available on CRAN, it would return
the canonical url form of the package 
(i.e. <https://CRAN.R-project.org/package=jsonlite>).

<details><summary>Example</summary>

```{r repository}

# Installed package

inst <- cff_create("jsonlite")

inst$repository

# Demo file downloaded from the r-universe

runiv <- system.file("examples/DESCRIPTION_r_universe", package = "cffr")
runiv_cff <- cff_create(runiv)

runiv_cff$repository

desc::desc(runiv)$get("Repository")

# For in development package

norepo <- system.file("examples/DESCRIPTION_basic", package = "cffr")

# No repo
norepo_cff <- cff_create(norepo)

norepo_cff[["repository"]]

# Change the name to a known package
# Create a temporary file
tmp <- tempfile("DESCRIPTION")
file.copy(norepo, tmp)


# Change name to a package on CRAN

desc::desc_set("Package", "ggplot2", file = tmp)

cff_create(tmp)[["repository"]]

# Show what happens if another repo is set

# Save original config
orig_options <- options()
getOption("repos")


# Set new repos
options(repos = c(
  tidyverse = "https://tidyverse.r-universe.dev",
  CRAN = "https://cloud.r-project.org"
))

# Load again the library
# Repos are evaluated on load
unloadNamespace("cffr")
library(cffr)


cff_create(tmp)[["repository"]]

# Now it is the tidyverse repo, due to our new config!

# Reset original config
options(orig_options)
getOption("repos")
```

</details>



### repository-artifact

This key is not extracted from the metadata of the package. See the
description on the [Guide
to CFF schema v1.2.0](https://github.com/citation-file-format/citation-file-format/blob/main/schema-guide.md#repository-artifact).

> - **description**: The URL of the work in a build artifact/binary repository 
> (when the work is software).
> - **usage**:<br><br>
>
>     ```yaml
>     repository-artifact: "https://search.maven.org/artifact/org.corpus-tools/cff-maven-plugin/0.4.0/maven-plugin"
>     ```


### repository-code

This key is extracted from the "BugReports" or "URL" fields on the 
DESCRIPTION file. **cffr** tries to identify the url of the source
on the following repositories:
 - [GitHub](https://github.com/).
 - [GitLab](https://about.gitlab.com/).
 - [R-Forge](https://r-forge.r-project.org/).
 - [Bitbucket](https://bitbucket.org/).


<details><summary>Example</summary>

```{r repository-code}

# Installed package on GitHub

cff_create("jsonlite")$`repository-code`



# GitLab

gitlab <- system.file("examples/DESCRIPTION_gitlab", package = "cffr")
cff_create(gitlab)$`repository-code`


# Check

desc::desc(gitlab)
```
</details>


### title

This key is extracted from the "Description" field of the DESCRIPTION file.


```{r eval=FALSE}
title <- paste0(
  "NAME_OF_THE_PACKAGE",
  ": ",
  "TITLE_OF_THE_PACKAGE"
)
```


<details><summary>Example</summary>

```{r title}

# Installed package

cff_create("testthat")$title
```
</details>


### type

Fixed value equal to "software". The other possible value is "dataset". See the
description on the [Guide
to CFF schema v1.2.0]https://github.com/citation-file-format/citation-file-format/blob/main/schema-guide.md#type).

### url

This key is extracted from the "BugReports" or "URL" fields on the 
DESCRIPTION file. It corresponds to the first url that is different to
[repository-code](#repository-code).

<details><summary>Example</summary>

```{r url}

# Many urls
manyurls <- system.file("examples/DESCRIPTION_many_urls", package = "cffr")

cff_create(manyurls)$url

# Check

desc::desc(manyurls)
```
</details>


### version

This key is extracted from the "Version" field on the 
DESCRIPTION file.


```{r version}

# Should be (>= 3.0.0)
cff_create("testthat")$version
```


## References
