Under the hood, a `cff` object is simply a named [`list`] to which we added
additional methods, most notably [`print`] and [as_cff()].

```{r}
#| include: false
library(cffr)
```

```{r}
a_named_list <- list(
  first = "I", second = "am", third = "a", fourth = "list",
  fifth = "with", sixth = "names", "none" = NULL
)

# Convert to `cff`.
a_cff_object <- as_cff(a_named_list)

class(a_cff_object)

a_cff_object

is.list(a_cff_object)
```

[as_cff()] not only converts a `list` to `cff` but also removes items that are
`NULL` or `NA`. These items are known as `keys` in CFF terminology.

### Subclasses

**cffr** implements two special subclasses of `cff`, which represent two special
types of objects defined in the [Citation File Format
schema](https://github.com/citation-file-format/citation-file-format/blob/main/schema-guide.md):

- `definitions.person` and `definitions.entity`: CFF definitions for sublists
  representing persons or entities. In **cffr**, the subclass `cff_pers_lst`
  collects an array of `definitions.person` or `definitions.entity` objects,
  where each individual person or entity has subclass `cff_pers`.

- Similarly, `definitions.reference` is the CFF definition for collecting
  references to related works and other software used in the development of the
  main work of the `CITATION.cff` file. In **cffr**, the subclass `cff_ref_lst`
  represents arrays of `definitions.reference` objects, where each element has
  subclass `cff_ref`.

These two subclasses do not provide full valid `cff` objects, but adapt
information to the schema required by CFF:

```{r}
# Use the utils::person() function.

## Array.
two_persons <- as_cff_person(
  c(
    person("A", "person", comment = c(ORCID = "0000-0001-8457-4658")),
    person("An entity", email = "fake@gmail.com")
  )
)

two_persons

class(two_persons)

# Single element.

two_persons[[1]]

class(two_persons[[1]])

# Array of references.

cit <- c(citation(), citation("jsonlite"))

ref_list <- as_cff(cit)

ref_list

class(ref_list)

# Single element.

ref_list[[1]]

class(ref_list[[1]])
```

## Valid `cff` objects

Creating a `cff` object does not ensure its validity according to the [Citation
File Format
schema](https://github.com/citation-file-format/citation-file-format/blob/main/schema-guide.md):

```{r}
class(a_cff_object)

cff_validate(a_cff_object)
```

[cff_validate()] gives minimal messages about what is wrong with the `cff`
object and invisibly returns the result of the validation (`TRUE/FALSE`).

We can use [cff_modify()] to add more keys:

```{r}
cff_valid <- cff_modify(a_cff_object,
  authors = as_cff_person("{James and James}"),
  cff_version = "1.2.0",
  message = "Hi there",
  title = "My title"
)


# Remove invalid keys.
cff_valid <- as_cff(cff_valid[names(cff_valid) %in% cff_schema_keys()])

cff_valid

cff_validate(cff_valid)
```

## Base methods provided by **cffr**

**cffr** version 1.0.0 provides additional S3 methods for common coercion
functions of the **base** and **utils** packages.

### [as.data.frame()]

```{r}
minimal_cff <- cff()

minimal_cff

as_df <- as.data.frame(minimal_cff)

class(as_df)

t(as_df)
```

### [c()]

```{r}
new_keys <- c("date-released" = "2020-01-31", abstract = "Minimal example")

c(minimal_cff, new_keys)
```

### [as.list()]

```{r}
as.list(minimal_cff)
```

### [as.person()]

Only for `cff_pers_lst` and `cff_pers` objects:

```{r}
as.person(two_persons)
```

### [toBibtex()]

```{r}
# For `cff`.
toBibtex(minimal_cff)

# For `cff_ref` and `cff_ref_lst`.
toBibtex(cit)

# For `cff_pers` and `cff_pers_lst`.
toBibtex(two_persons)
```
