## The `cff` class

[**cffr**](https://cran.r-project.org/package=cffr) implements a S3 object with
[`class`] `cff`, that it is used to represent the information of a
`CITATION.cff` file in **R**.

Under the hood, a `cff` object is simply a named [`list`] to which we added
additional methods, most notably [`print`] and [as_cff()].

```{r include=FALSE}
library(cffr)
```

```{r}
a_named_list <- list(
  first = "I", second = "am", third = "a", fourth = "list",
  fifth = "with", sixth = "names", "none" = NULL
)

# As cff
a_cff_object <- as_cff(a_named_list)

class(a_cff_object)

a_cff_object

is.list(a_cff_object)
```

[as_cff()] not only converts a `list` to `cff` but also removes items (known as
`keys` in CFF terminology) that are `NULL` or `NA`.

### Sub-classes

**cffr** implements two special sub-classes of `cff`, with the aim of
representing two special types of objects defined in the [CFF
Schema](https://github.com/citation-file-format/citation-file-format/blob/main/schema-guide.md):

-   `definition.person` and `definition.entity`: CFF definition for sub-lists
    representing persons or entities. In **cffr** the sub-class `cff_pers_lst`
    has been implemented to collect an array of `definition.person/entity`,
    where each individual person or entity of the array has a sub-class
    `cff_pers`.

-   Similarly, `definition.reference` is the definition of CFF for collecting
    references to related works and other articles of software used in the
    development of the main work of the CFF file. This is implemented in
    **cffr** as a sub-class named `cff_ref_lst` for arrays of
    `definition.reference` where each element of the array has a sub-class named
    `cff_ref`.

These two sub-classes does not provide full valid `cff` objects, but adapts
information to the schema required by CFF:

```{r}
# Using the utils::person() function

## Array
two_persons <- as_cff_person(
  c(
    person("A", "person", comment = c(ORCID = "0000-0000-0000-0000")),
    person("An entity", email = "fake@gmail.com")
  )
)

two_persons

class(two_persons)

# Single element

two_persons[[1]]

class(two_persons[[1]])

# Array of references

cit <- c(citation(), citation("yaml"))

ref_list <- as_cff(cit)

ref_list

class(ref_list)

# Single element

ref_list[[1]]

class(ref_list[[1]])
```

## Valid `cff` objects

Creating a `cff` object does not ensure its validity according to the [CFF
Schema](https://github.com/citation-file-format/citation-file-format/blob/main/schema-guide.md):

```{r}
class(a_cff_object)

cff_validate(a_cff_object)
```

[cff_validate()] gives minimal messages of what's wrong with our `cff` and
(invisibly) returns the result of the validation (`TRUE/FALSE`).

We can use [cff_modify()] to add more keys:

```{r}
cff_valid <- cff_modify(a_cff_object,
  authors = as_cff_person("{James and James}"),
  cff_version = "1.2.0",
  message = "Hi there",
  title = "My title"
)


# Remove invalid keys
cff_valid <- as_cff(cff_valid[names(cff_valid) %in% cff_schema_keys()])

cff_valid

cff_validate(cff_valid)
```

## Base methods provided by **cffr**

**cffr** version 1.0.0 provides additional S3 Methods for common coercing
functions of the **base** and **utils** package.

### [as.data.frame()]

```{r}
minimal_cff <- cff()

minimal_cff

as_df <- as.data.frame(minimal_cff)

class(as_df)

t(as_df)
```

### [c()]

```{r}
new_keys <- c("date-released" = "2020-01-31", abstract = "Minimal example")

c(minimal_cff, new_keys)
```

### [as.list()]

```{r}
as.list(minimal_cff)
```

### [as.person()]

Only for `cff_pers_lst` and `cff_pers` objects:

```{r}
as.person(two_persons)
```

### [toBibtex()]

```{r}
# For cff
toBibtex(minimal_cff)

# cff_ref, cff_ref_lst
toBibtex(cit)

# cff_pers, cff_pers_lst
toBibtex(two_persons)
```
