## The `cff` class

**cffr** implements a S3 object with [class()] `cff`, that it is used to
represent the information of a `*.cff` file in **R**.

Under the hood, a `cff` object is simply a named [list()] to which we added
additional methods, most notably [print()].

```{r}
a_named_list <- list(
  first = "I", second = "am", third = "a", fourth = "list",
  fifth = "with", sixth = "names", "none" = NULL
)


dput(a_named_list)

# Default print
a_named_list

# But
a_cff_object <- as.cff(a_named_list)

class(a_cff_object)

a_cff_object

dput(a_cff_object)
```

[as.cff()] not only converts a [list()] to `cff` but also removes items (known
as `keys` in CFF terminology) that are `NULL` or `NA`.

## Valid `cff` objects

Although `a_cff_object` is a `cff` object, it is not valid, since it does not
comply with the validation rules of the CFF specification.

```{r}
cff_validate(a_cff_object)
```

[cff_validate()] gives minimal messages of what's wrong with our `cff` and
(invisibly) returns the result of the validation (`TRUE/FALSE`).

## Other methods

### [as.data.frame()]

```{r}
minimal_cff <- cff()

minimal_cff

as_df <- as.data.frame(minimal_cff)

class(as_df)

t(as_df)
```

## [c()]

```{r}
new_keys <- c("date-released" = "2020-01-31", abstract = "Minimal example")

c(minimal_cff, new_keys)
```

## [as.person()]

Special case for those CFF keys that are person-like.

```{r}
path <- system.file("examples/CITATION_complete.cff", package = "cffr")
the_cff <- cff_read(path)

the_cff$authors

as.person(the_cff$authors)
```
